---
title: "Running TSNE on any stock"
author: "Maxandre Hebert"
date: "2023-05-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Quick remember
TSNE is used to find stock that move in a similar way.(Tend to move together)
To perform a t-sne on stock we need to put each date as a feature(column).
That took me a long time to do, but it was worth it.

## load library
```{r}
library(tidyverse)
library(Rtsne)
library(plotly) # for the plot
library(yfR) #getting stock data
```

## Getting the data

```{r}
#Here is just an exemple of ticker
tickers <- c("AAPL", "GOOGL", "MSFT", "AMZN", "FB", "TSLA", "NVDA", "JNJ", "V", "PG")
df_sp500 <- yf_index_composition("SP500") %>% filter(ticker %in% tickers)
  my_df <- yf_get(
    tickers,
    first_date = "2007-01-01",
    last_date = Sys.Date(),
    thresh_bad_data = FALSE, 
    bench_ticker = "^GSPC",
    type_return = "arit",
    freq_data = "daily",
    how_to_aggregate = "last",
    do_complete_data = FALSE,
    do_cache = TRUE,
    do_parallel = FALSE,
    be_quiet = FALSE
  )
```

##Data wrangling
1. We need to find the minimum starting date for stock.
2. We then put each date as features, to do that we use function pivot_wider
from the package tidyr.
-Def : "widens" data, increasing the number of columns and decreasing the number of rows. The inverse transformation is "pivot_longer"
```{r}
# Step 1  : Calculate the first date for each stock
start_dates <- function() {
  x <- my_df %>%
  group_by(ticker) %>%
  summarise(start_date = min(ref_date))
  
  return(max(x$start_date))
}

# Step 2 : filter the data
filtered_df <- my_df %>% group_by(ticker) %>%
  select(ticker,ref_date,ret_adjusted_prices) %>%
  filter(ref_date > start_dates()) %>%
  pivot_wider(names_from = ref_date, values_from = ret_adjusted_prices)

```

##Deploy the t-sne
Now we can deploy the t-sne.
What is t-SNE? t-Distributed Stochastic Neighbor Embedding (t-SNE) is an unsupervised, non-linear technique primarily used for data exploration and visualizing high-dimensional data
```{r}
tsne_data <- filtered_df[, -1]  # assuming the first column is `ticker`

tsne_model <- Rtsne(tsne_data, dims = 2, perplexity=1, theta=0.5, check_duplicates=FALSE)

# the left join function is mainly for adding sector to their ticker names
tsne_results <- data.frame(ticker = filtered_df$ticker, tsne_model$Y) %>%
  left_join(df_sp500 %>% select(ticker, sector), by = "ticker")



```

##Data explatory
```{r}
p <- ggplot(tsne_results, aes(x = X1, y = X2, text=ticker,color=sector)) +
  geom_point() +
  geom_text(aes(label=sector), size=2, vjust=1, hjust=0.5,nudge_y = 0.2) +
  theme_minimal() +
  ggtitle("t-SNE plot of Stocks")

p <- ggplotly(p)

p

```

## Bringing all together
```{r}
tsne_stock_map <- function(ticker_vector) {
  require(tidyverse)
  require(Rtsne)
  require(plotly) # for the plot
  require(yfR) #getting stock data
  #Here is just an exemple of ticker
  
  tickers <- ticker_vector
  df_sp500 <- yf_index_composition("SP500") %>% filter(ticker %in% tickers)
  my_df <- yf_get(
      tickers,
      first_date = "2007-01-01",
      last_date = Sys.Date(),
      thresh_bad_data = FALSE, 
      bench_ticker = "^GSPC",
      type_return = "arit",
      freq_data = "daily",
      how_to_aggregate = "last",
      do_complete_data = FALSE,
      do_cache = TRUE,
      do_parallel = FALSE,
      be_quiet = FALSE
    )
  # Step 1  : Calculate the first date for each stock
  start_dates <- my_df %>%
    group_by(ticker) %>%
    summarise(start_date = min(ref_date)) %>%
    max()
  
  # Step 2 : filter the data
  filtered_df <- my_df %>% group_by(ticker) %>%
    select(ticker,ref_date,ret_adjusted_prices) %>%
    filter(ref_date > start_dates) %>%
    pivot_wider(names_from = ref_date, values_from = ret_adjusted_prices)
  
  tsne_data <- filtered_df[, -1]  # assuming the first column is `ticker`

  tsne_model <- Rtsne(tsne_data, 
                      dims = 2, 
                      perplexity=30, 
                      theta=0.5, 
                      check_duplicates=FALSE)
  
  # the left join function is mainly for adding sector to their ticker names
  tsne_results <- data.frame(ticker = filtered_df$ticker, tsne_model$Y) %>%
    left_join(test %>% select(ticker, sector), by = "ticker")
  
  p <- ggplot(tsne_results, 
              aes(x = X1, 
                  y = X2, 
                  text=ticker,color=sector)) +
  geom_point() +
  geom_text(aes(label=sector), 
            size=2, 
            vjust=1, 
            hjust=0.5,
            nudge_y = 0.2) +
  theme_minimal() +
  ggtitle("t-SNE plot of Stocks")

  p <- ggplotly(p)
  
  p
    
  
  return(p,tsne_results)
}


```